### 2017-10-24 Draft2/Abstract.tex
Revise Intro, Abstract

This paper introduces the \textbf{growth array}, an array-like data structure designed to be efficient at appending elements. They offer substantial time and memory savings over dynamic arrays when a large number of elements are appended. Growth arrays also support other list operations, such as random access, iteration, and insertion/deletion at an index. However, they perform no better, or slightly worse, than dynamic arrays for those methods. Also, if the size of the data is known beforehand, allocating a raw array with that size and appending to it is faster than appending to a growth array.

>>>>

This paper introduces \textbf{growth arrays}, array-like data structures that are designed for appending elements. When the number of items is not known beforehand but is expected to be large, they are more efficient than dynamic arrays by a constant factor. Growth arrays support all operations dynamic arrays do, such as random access, iteration, and insertion or deletion at an index. However, they perform no better, or slightly worse, than dynamic arrays for operations other than appending.

### 2017-10-24 Draft2/Intro.tex
Remove footnotes
Revise Intro, Abstract

In imperative languages, the dynamic array is the most common data structure used by programs\footnote{}. People often want to add multiple items to a list and then iterate it, which dynamic arrays make simple and efficient. But can it be said dynamic arrays have the \textit{most} efficient algorithm for this pattern?

In this paper, I introduce an alternative data structure to the dynamic array, called the \textbf{growth array}. It is more efficient than the dynamic array at appending, particularly for large numbers of items. This is due to how it 'grows' once it cannot fit any more items in its buffer.

When dynamic arrays run out of space, they allocate a new buffer, copy the contents of the old one into it, and throw the old one away. However, growth arrays are less wasteful. Instead of throwing away the filled buffer, they keep it a part of the data structure. The new buffer they allocate represents a continuation of the items from the old buffer. For example, if the old buffer contained items $0-31$, the new buffer would contain items $32$ and beyond. Because of this, growth arrays allocate less memory to store the same number of items, and they do not need to copy items from the old buffer to the new one.

Growth arrays have some caveats, however. They perform no better, or slightly worse, than dynamic arrays with respect to other operations. In particular, random access involves significantly more instructions than it does for dynamic arrays. Additionally, since growth arrays are not contiguous in memory, they may have poorer locality than dynamic arrays, and cannot be passed to external code that accepts contiguous buffers.

It is also worth mentioning that if the size of the data is known in advance, both dynamic and growth arrays are completely unnecessary. A raw array could simply be allocated with the known size, and items could be appended more quickly to it. Thus, growth arrays are only advantageous for scenarios where an unknown amount of data will be appended.

>>>>

In imperative languages, the dynamic array is the most common data structure used by programs. People often want to add multiple items to a collection and then iterate it, which dynamic arrays make simple and efficient. But can it be said they have the \textit{most} efficient algorithm for this pattern?

In this paper, I introduce an alternative data structure to the dynamic array, called the \textbf{growth array}. It is more efficient than the dynamic array at appending large numbers of items. This is due to how it 'grows' once it cannot fit more items in its buffer.

When dynamic arrays run out of space, they allocate a new buffer, copy the contents of the old one into it, and throw the old one away. However, growth arrays are less wasteful. Instead of throwing away the filled buffer, they keep it a part of the data structure. The new buffer they allocate represents a continuation of the items from the old buffer. For example, if the old buffer contained items $0 - 31$, the new buffer would contain items $32$ and beyond. Because of this, growth arrays allocate less memory to store the same number of items, and they do not need to copy items from the old buffer to the new one.

Growth arrays have caveats, however. They perform no better, or slightly worse, than dynamic arrays for operations other than appending. Random access, in particular, involves many more instructions than it does for dynamic arrays. Also, since growth arrays are not contiguous in memory, they may have poorer locality than dynamic arrays, and cannot be passed to external code that accepts contiguous buffers.

It is worth mentioning that if the size of the data is known in advance, both dynamic and growth arrays are completely unnecessary. A raw array could simply be allocated with the known size, and items could be appended to it just as quickly. Thus, growth arrays are only beneficial for cases where the amount of data to be appended is unknown, but is expected to be large.
